Leadership across cultures and contexts: an exploratory analysis

### 2020-04-30 leadership_across_cultures_contexts.Rmd
Minor revisions to new intro section. Added references.

<!-- INTRO TO MOTIVATE PAPER -->
Research on leadership within anthropology dates to the inception of the discipline. Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917]. Anthropologists documented some sort of leadership among every ethnographically studied culture [@brown_human_1991; @lewis_leaders_1974], and in many social contexts, such as families and kin groups (Dussart, 2000), ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], and conflicts between groups [@glowacki_evolutionary_2017]. For review, see @garfield_review_2019.

Despite decades of research and theorizing about leadership in numerous social science disciplines, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of cultures. Moreover, outside of anthropology, most research on leadership is based on data from Western or Westernized societies.

@@

There is now well established consensus within behavioral and evolutionary human sciences on the importance of data from the full range of human cultural diversity. Systemic biases in much of social science are well known [@henrich_weirdest_2010, *The weirdest people in the world?*], and might be due, in part, to the lack of data from non-WEIRD populations. @henrich_weirdest_2010 in their initial critique of the over-realiance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.

@@

>culture is important to anthropology because the anthropological distinction between self and other rests on it. Culture is the essential tool for making other. As a professional discourse that elaborates on the meaning of culture in order to account for, explain, and understand cultural difference,anthropology also helps construct, produce, and maintain it. Anthropological discourse gives cultural difference (and the separation between groups of people it implies) the air of the self-evident. In this regard, the concept of culture operates much like its predecessor - race - even though in its twentieth-century form it has some important political advantages. Unlike race, and unlike even the nineteenth-century sense of culture as a synonym for civilization (contrasted to barbarism), the current concept allows for multiple rather than binary differences.

@@

<!-- EVOLUTIONARY THEORY APPROACH -->
Much scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover the interrelationships of these dimensions of leadership. Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology. Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @redhead_dynamics_2018; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018; @lonati_what_2019]. This integration as been informative [see @van_vugt_genes_2020].

>>>>

<!-- INTRO TO MOTIVATE PAPER -->
Research on leadership within anthropology extends to the birth of the discipline. Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917]. Anthropologists documented some sort of leadership among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and across many social contexts, such as families and kin groups [@dussart2000politics], ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967]. For review, see @garfield_review_2019.

Despite decades of research and theorizing about leadership in numerous social science disciplines, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies. Moreover, outside of anthropology, most research and theorizing on leadership is based on data from Western or Westernized, postindustrial societies [@von_rueden_leadership_2015].

@@

There is now well established consensus within the behavioral sciences on the importance of data from the full range of human cultural diversity. Systemic biases in much of social science are well known [@henrich_weirdest_2010, *The weirdest people in the world?*], and might be due, in part, to the lack of data from populations other than American undergaduates. @henrich_weirdest_2010 in their initial critique of the over-realiance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.

@@

>culture is important to anthropology because the anthropological distinction between self and other rests on it. Culture is the essential tool for making other. As a professional discourse that elaborates on the meaning of culture in order to account for, explain, and understand cultural difference, anthropology also helps construct, produce, and maintain it. Anthropological discourse gives cultural difference (and the separation between groups of people it implies) the air of the self-evident. In this regard, the concept of culture operates much like its predecessor - race - even though in its twentieth-century form it has some important political advantages. Unlike race, and unlike even the nineteenth-century sense of culture as a synonym for civilization (contrasted to barbarism), the current concept allows for multiple rather than binary differences.

@@

<!-- EVOLUTIONARY THEORY APPROACH -->
Much scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover the interrelationships of these dimensions of leadership. Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology. Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @wiessner_collective_2019; @redhead_dynamics_2018; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018; @lonati_what_2019]. This integration as been informative [see @van_vugt_genes_2020].
